---
title: "ECON320Lec_HW3"
author: "D.Z."
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(ggplot2)
library(stargazer)
library(kableExtra)
library(foreign)
```

## Part 1

i. Generate a dataset of 100,000 observations from this DGP. Call this the "population data"

```{r echo = TRUE}
x <- rnorm(100000, 10, 10)
u <- rnorm(100000, 0, 5)
y = 2.5 + 1.5 * x + u
population_data <- data.frame(x, u, y)
```

ii. Draw a random sample of 1,000 observations from the population data. Estimate the parameters $\beta_0$ and $\beta_1$.
```{r echo = TRUE}
randX1000 <- sample_n(population_data, 1000)
randX1000Reg <- lm(formula = y ~ x, data = randX1000)
summary(randX1000Reg)
(beta_0 <- summary(randX1000Reg)$coefficients[1, 1])
(beta_1 <- summary(randX1000Reg)$coefficients[2, 1])
```

iii. Repeat step ii. 1000 times, and save the estimated parameters from each draw
```{r echo = TRUE}
beta_0_vector <- c()
beta_1_vector <- c()
for(i in 1:1000) {
  newPop <- population_data[sample(nrow(population_data), 1000), ]
  beta_0 <- mean(newPop$y) - mean(newPop$x) * beta_1
  beta_1 <- cov(newPop$y, newPop$x) / var(newPop$x)
  beta_0_vector <- c(beta_0_vector, beta_0)
  beta_1_vector <- c(beta_1_vector, beta_1)
}
```

iv. Present histograms of the sampling distribution of the estimates of $\beta_0$ and $\beta_1$. Call these Figures 1 and 2 (with appropriate titles).
```{r echo = TRUE}
hist(beta_0_vector, main = "Figure 1", xlab = "beta_0", breaks = 30)
hist(beta_1_vector, main = "Figure 2", xlab = "beta_1", breaks = 30)
```

v. What happens to the sampling distributions of the estimate of $\beta_1$ if U ~ N(0, 50) or U ~ N(0, 1)? Present two more histograms, Figures 3 and 4.
```{r echo = TRUE}
x1 <- rnorm(100000, 10, 10)
u1 <- rnorm(100000, 0, 50)
y1 = 2.5 + 1.5 * x + u
situation_1 <- data.frame(x1, u1, y1)
beta_1_vector1 <- c()
for(i in 1:1000) {
  newPop1 <- situation_1[sample(nrow(situation_1), 1000), ]
  beta_1_1 <- cov(newPop1$y, newPop1$x) / var(newPop1$x)
  beta_1_vector1 <- c(beta_1_vector1, beta_1_1)
}
hist(beta_1_vector1, main = "Figure 3", xlab = "beta_1_1", breaks = 30)

x2 <- rnorm(100000, 10, 10)
u2 <- rnorm(100000, 0, 1)
y2 = 2.5 + 1.5 * x + u
situation_2 <- data.frame(x2, u2, y2)
beta_1_vector2 <- c()
for(i in 1:1000) {
  newPop2 <- situation_2[sample(nrow(situation_2), 1000), ]
  beta_1_2 <- cov(newPop2$y, newPop2$x) / var(newPop2$x)
  beta_1_vector2 <- c(beta_1_vector2, beta_1_2)
}
hist(beta_1_vector2, main = "Figure 4", xlab = "beta_1_2", breaks = 30)
```

## Part 2. a

a.
```{r echo = TRUE}
card_returns <- read.dta("/Users/dingyuanzhang/Documents/F19_Eco320L/320Lec/Card_Returns.dta")
card_returns$married[card_returns$married>1]<-0

card_returns %>% select(wage, educ, exper) %>% summarise(wage_mean = mean(wage), wage_sd = sd(wage), wage_min = min(wage), wage_max = max(wage), educ_mean = mean(educ), educ_sd = sd(educ), educ_min = min(educ), educ_max = max(educ), exper_mean = mean(exper), exper_sd = sd(exper), exper_min = min(exper), exper_max = max(exper)) %>% gather(value) %>% setNames(., c("Table 1", ""))%>% kable() %>% kable_styling() %>% add_footnote(paste0("observations count: ", NROW(card_returns), " ", "variables count: ", NCOL(card_returns)), notation= "number")

plot<-ggplot(card_returns, aes(x = educ, y = wage))+ 
  geom_point(color="red")+
  geom_smooth(method='lm') +
  ggtitle(paste("Figure 5")) 
print(plot)
```

b. 
```{r echo = TRUE}
spec1 <- lm(formula = log(wage) ~ educ, data = card_returns)
summary(spec1)

spec2 <- lm(formula = log(wage) ~ educ + exper, data = card_returns)
summary(spec2)

spec3 <- lm(formula = log(wage) ~ educ + exper + expersq, data = card_returns)
summary(spec3)

spec4 <- lm(formula = log(wage) ~ educ + exper + expersq + married, data = card_returns)
summary(spec4)

spec5 <- lm(formula = log(wage) ~ educ + exper + expersq + married + black, data = card_returns)
summary(spec5)
```
Table 2

Variables | Spec1 | Spec2 | Spec3 | Spec4 | Spec5
----------|-------|-------|-------|-------|------
$\beta_0$ |`r summary(spec1)$coefficients[1, 1]`|`r summary(spec2)$coefficients[1, 1]`|`r summary(spec3)$coefficients[1, 1]`|`r summary(spec4)$coefficients[1, 1]`|`r summary(spec5)$coefficients[1, 1]`
$\beta_1$ |`r summary(spec1)$coefficients[2, 1]`|`r summary(spec2)$coefficients[2, 1]`|`r summary(spec3)$coefficients[2, 1]`|`r summary(spec4)$coefficients[2, 1]`|`r summary(spec5)$coefficients[2, 1]`
$\beta_2$ ||`r summary(spec2)$coefficients[3, 1]`|`r summary(spec3)$coefficients[3, 1]`|`r summary(spec4)$coefficients[3, 1]`|`r summary(spec5)$coefficients[3, 1]`
$\beta_3$ |||`r summary(spec3)$coefficients[4, 1]`|`r summary(spec4)$coefficients[4, 1]`|`r summary(spec5)$coefficients[4, 1]`
$\beta_4$ ||||`r summary(spec4)$coefficients[5, 1]`|`r summary(spec5)$coefficients[5, 1]`
$\beta_5$ |||||`r summary(spec5)$coefficients[6, 1]`

The value of $\beta_0$ in specification 2 is `r summary(spec2)$coefficients[1, 1]`, and that in specification 5 is `r summary(spec5)$coefficients[1, 1]`. The proportion of difference between these two variables is `r ((summary(spec2)$coefficients[1, 1] - summary(spec5)$coefficients[1, 1]) / summary(spec2)$coefficients[1, 1]) * 100`%. The value of $\beta_1$ in specification 2 is `r summary(spec2)$coefficients[2, 1]`, and that in specification 5 is `r summary(spec5)$coefficients[2, 1]`. The proportion of difference between these two values is `r ((summary(spec2)$coefficients[2, 1] - summary(spec5)$coefficients[2, 1]) / summary(spec2)$coefficients[2, 1]) * 100`%. The value of $\beta_2$ in specification 2 is `r summary(spec2)$coefficients[3, 1]`, and that in specification 5 is `r summary(spec5)$coefficients[3, 1]`. The proportion of increase between these two variables is `r ((summary(spec2)$coefficients[3, 1] - summary(spec5)$coefficients[3, 1]) / summary(spec2)$coefficients[3, 1]) * 100`%. The change in $\beta_2$ is very siginificant.

c. 

In specification 2, if a worker gains one more year of experience, the increase in wage is e^`r summary(spec2)$coefficients[3, 1]` unit. In specification 3, if a worker gains one more year of experience, the increase in wage is (e^`r summary(spec3)$coefficients[3, 1]` + (e^(`r summary(spec3)$coefficients[4, 1]`*exper))) unit.

d.
```{r echo = TRUE}
y_hat_1 <- fitted(spec1)
y_hat_5 <- fitted(spec5)
res_1 <- log(card_returns$wage)-y_hat_1
res_5 <- log(card_returns$wage)-y_hat_5
(sum(res_1))
(sum(res_5))
```

Properties of OLS estimatior (1 and 2):
  1. The estimators for each indipendent variables are linear.
  2. The OLS estimators are unbiased.
In specification 1, all parameters are linear, therefore OLS property 1 holds.
In specification 5, all parameters are linear as well, therefore OLS property 1 holds.
The residual in specification 1 is `r sum(res_1)`, and the residual in specification 5 is `r sum(res_1)`. Neither residuals is zero. The one of specification 1 is very close to zero, but the one of specificaiton 5, compared to that in specification 1, is not very close to zero. Therefore, the estimators for specification 5 is biased.


## Part 2. b

a.
```{r echo = TRUE}
HIV <- read.dta("/Users/dingyuanzhang/Documents/F19_Eco320L/320Lec/Thornton_Demand for HIV Status.dta")
HIV %>% select(got, any, tinc, male, hiv2004) %>% na.omit() %>% summarise(got_mean = mean(got), got_sd = sd(got), got_min = min(got), got_max = max(got), any_mean = mean(any), any_sd = sd(any), any_min = min(any), any_max = max(any), tinc_mean = mean(tinc), tinc_sd = sd(tinc), tinc_min = min(any), tinc_max = max(any), male_mean = mean(male), male_sd = sd(male), male_min = min(male), male_max = max(male), hiv2004_mean = mean(hiv2004), hiv2004_sd = sd(hiv2004), hiv2004_min = min(hiv2004), hiv2004_max = max(hiv2004)) %>% gather(value) %>% setNames(., c("Table 3", ""))%>% kable() %>% kable_styling() %>% add_footnote(paste0("observations count: ", NROW(HIV), " ", "variables count: ", NCOL(HIV)), notation= "number")
```

b.
```{r echo = TRUE, results = "asis"}
# lm(formula = got ~ any + (0.075 * tinc) + I((0.075 * tinc)^2) + distvct + I(distvct^2) + cov(male, age, age2, hiv2004, site, simaverage), data = HIV_omit)
HIV1 <- lm(formula = got ~ any + hiv2004 + male + age + rumphi + balaka, data = HIV)
HIV2 <- lm(formula = got ~ any + I(0.075*tinc) + hiv2004 + male + age + rumphi + balaka, data = HIV)
HIV3 <- lm(formula = got ~ any + I(0.075*tinc) + I((0.075 * tinc)^2) + hiv2004 + male + age + rumphi + balaka, data = HIV)
stargazer(HIV1, HIV2, HIV3, title = "Table 4", align = TRUE, covariate.labels = c("Any incentive", "Amount of incentive", "HIV", "Distance(km)", "Distance square", "Over 1.5 km", "Male", "Age", "Simulated", "Rumphi", "Balaka"), type = "html")
```

## Part 2. c

a.
```{r echo = TRUE}
lakisha_aer <- read.dta("/Users/dingyuanzhang/Documents/F19_Eco320L/320Lec/BertrandMullainathan_Emily and Greg.dta")

lakisha_aer %>% select(call, h, race, sex, ofjobs, yearsexp) %>% summarise(call_mean = mean(call), call_sd = sd(call), call_min = min(call), call_max = max(call), h_mean = mean(h), h_sd = sd(h), h_min = min(h), h_max = max(h), race_mean = mean(race), race_sd = sd(race), race_min = min(race), race_max = max(race), sex_mean = mean(sex), sex_sd = sd(sex), sex_min = min(sex), sex_max = max(sex), ofjobs_mean = mean(ofjobs), ofjobs_sd = sd(ofjobs), ofjobs_min = min(ofjobs), ofjobs_max = max(ofjobs), yearsexp_mean = mean(yearsexp), yearsexp_sd = sd(yearsexp), yearsexp_min = min(yearsexp), yearsexp_max = max(yearsexp)) %>% gather(value) %>% setNames(., c("Table 5", ""))%>% kable() %>% kable_styling() %>% add_footnote(paste0("observations count: ", NROW(lakisha_aer), " ", "variables count: ", NCOL(lakisha_aer)), notation= "number")
```

b.
```{r echo = TRUE, results = "asis"}
white_names <- lakisha_aer %>% filter(race == 0)
African_American_names <- lakisha_aer %>% filter(race == 1)
lakisha1 <- lm(formula = call ~ I(yearsexp/10) + I(yearsexp^2/100) + volunteer + military + email + empholes + workinschool + honors + computerskills + specialskills, data = lakisha_aer)
lakisha2 <- lm(formula = call ~ I(yearsexp/10) + I(yearsexp^2/100) + volunteer + military + email + empholes + workinschool + honors + computerskills + specialskills, data = white_names)
lakisha3 <- lm(formula = call ~ I(yearsexp/10) + I(yearsexp^2/100) + volunteer + military + email + empholes + workinschool + honors + computerskills + specialskills, data = African_American_names)
stargazer(lakisha1, lakisha2, lakisha3, title = "Table 5", align = TRUE,column.lables = c("All Resumes", "White Names", "African American Names"), covariate.labels = c("Years of experience (10)", "Years of experience squared (100)", "Volunteering? (Y = 1)", "Military experience? (Y = 1)", "Email? (Y = 1)", "Employment holes? (Y = 1)", "Computer skills? (Y = 1)", "Special skills? (Y = 1)"), type = 'html')
```

